The Power of Unbiased Recursive Partitioning:\\ A Unifying View of CTree, MOB, and GUIDE

### 2019-03-06 pkg/partykit/inst/guideliketest/DAGStat2019/slides.Rnw
revision of recommendations and further small corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2484 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 02da08f193b42d809a6aa3269f6ff390613395a9

## HCL palette
pal_dgp <- qualitative_hcl(3, "Set 2")

@@

{\bf General testing strategy:}
\begin{enumerate}
\item Evaluate a discrepancy measure capturing the obervation-wise goodness of fit of $\mathcal{M}(Y,X;\hat{\beta})$.
\item Apply a statistical test assessing dependency of the discrepancy measure to each possible split variable $Z_j$.
\item Select the split variable $Z_j^\ast$ showing the smallest $p$-value.
\end{enumerate}

@@

\begin{frame}
\frametitle{A Unifying view}

The testing strategies of CTree, MOB, and GUIDE can be obtained from the general 
framework by specific combinations of the building blocks together with the 
corresponding approximation type of the underlying null distribution.

@@

\begin{table}[t]
\begin{tabular}{lllll}
\hline 
      & Scores       & Binarization  & Categorization  & Statistic \\ 
\hline
CTree & Model scores & --	   & -- 	   & Sum of squares\\
MOB   & Model scores & --	   & -- 	   & Maximally selected\\
GUIDE & Residuals    & \checkmark  & \checkmark    & Sum of squares\\
\hline
\end{tabular}
\end{table}
\end{frame}

@@

\begin{itemize}
\item {\only<4->{\color{lightgray}} Residuals vs.\ full model scores}
\item {\only<1-3,6->{\color{lightgray}} Binarization of residuals/scores}
\item {\only<1-5>{\color{lightgray}} Categorization of possible split variables}
\end{itemize}

@@

\only<2>{
\small
$$
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) \cdot X}\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) \cdot X}\\
\vdots & {\color{lightgray} \vdots} \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) \cdot X}
\end{pmatrix}
%\begin{pmatrix}
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})}\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})}\\
%\vdots & {\color{lightgray} \vdots} \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})}
%\end{pmatrix}
$$
}
\only<3>{
\small
$$
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) \cdot X\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) \cdot X\\
\vdots & \vdots \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) \cdot X
\end{pmatrix}
%\begin{pmatrix} 
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
%\vdots & \vdots \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
%\end{pmatrix}
$$
}
\only<4-5>{
\small
$$
r(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
\begin{pmatrix}
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
\vdots \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
\end{pmatrix}
%
\visible<5>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
> 0\\
\le 0\\
\vdots \\
> 0
\end{pmatrix}
}
$$
}
\only<6-7>{
\small
$$
Z_j = 
\begin{pmatrix}
Z_{j1}\\
Z_{j2}\\
\vdots \\
Z_{jn}\\
\end{pmatrix}
%
\visible<7>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
\text{Q}3\\
\text{Q}1\\
\vdots \\
\text{Q}2
\end{pmatrix}
}
$$
}
\end{frame}

@@

tp <- function (obj, digits = 3, abbreviate = FALSE, 
                fill = c("lightgray", "white"), id = TRUE, 
                just = c("center", "top"), top = 0.85, 
                align = c("center", "left", "right"), gp = NULL, FUN = NULL, 
                height = NULL, width = NULL) 
{
  nam <- names(obj)
  extract_label <- function(node) formatinfo_node(node, FUN = FUN, 
                                                  default = c("terminal", "node"))
  maxstr <- function(node) {
    lab <- extract_label(node)
    klab <- if (is.terminal(node)) 
      ""
    else unlist(lapply(kids_node(node), maxstr))
    lab <- c(lab, klab)
    lab <- try(unlist(lapply(lab, function(x) strsplit(x, 
                                                       "\n"))), silent = TRUE)
    if (inherits(lab, "try-error")) {
      paste(rep("a", 9L), collapse = "")
    }
    else {
      return(lab[which.max(nchar(lab))])
    }
  }
  nstr <- if (is.null(width)) 
    maxstr(node_party(obj))
  else paste(rep("a", width), collapse = "")
  just <- match.arg(just[1L], c("center", "centre", "top"))
  if (just == "centre") 
    just <- "center"
  align <- match.arg(align[1L], c("center", "centre", "left", 
                                  "right"))
  if (align == "centre") 
    align <- "center"
  rval <- function(node) {
    if(node$id == 2) fill <- c("white", pal_dgp[1])
    if(node$id == 3 | node$id == 4) fill <- c("white", pal_dgp[2])
    if(node$id == 5) fill <- c("white", pal_dgp[3])
    fill <- rep(fill, length.out = 2)
    lab <- extract_label(node)
    if (!is.null(gp)) {
      outer_vp <- viewport(gp = gp)
      pushViewport(outer_vp)
    }
    if (is.null(height)) 
      height <- length(lab) + 1L
    node_vp <- viewport(x = unit(0.5, "npc"), y = unit(if (just == 
                                                           "top") 
      top
      else 0.5, "npc"), just = c("center", just), width = unit(1, 
                                                               "strwidth", nstr) * 1.1, height = unit(height, "lines"), 
      name = paste("node_terminal", id_node(node), sep = ""), 
      gp = if (is.null(gp)) 
        gpar()
      else gp)
    pushViewport(node_vp)
    grid.rect(gp = gpar(fill = fill[1]))
    for (i in seq_along(lab)) grid.text(x = switch(align, 
                                                   center = unit(0.5, "npc"), left = unit(1, "strwidth", 
                                                                                          "a"), right = unit(1, "npc") - unit(1, "strwidth", 
                                                                                                                              "a")), y = unit(length(lab) - i + 1, "lines"), 
                                        lab[i], just = align)
    if (id) {
      nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                                                          "npc"), width = max(unit(1, "lines"), unit(1.3, 
                                                                                                     "strwidth", nam[id_node(node)])), height = max(unit(1, 
                                                                                                                                                         "lines"), unit(1.3, "strheight", nam[id_node(node)])))
      pushViewport(nodeIDvp)
      grid.rect(gp = gpar(fill = fill[2], lty = "solid"))
      grid.text(nam[id_node(node)])
      popViewport()
    }
    if (is.null(gp)) 
      upViewport()
    else upViewport(2)
  }
  return(rval)
}

@@

               strip =strip.custom(factor.levels = c(expression("vary"~beta[0]~"and"~beta[1]), 
                                                     expression("vary"~beta[0]), 
                                                     expression("vary"~beta[1])),
                                   bg = "lightgray"),
               strip.left=strip.custom(factor.levels = c(expression(xi == 0),                                                              expression(xi == 0.8)),
                                       bg = "lightgray"))

@@

\begin{frame}%[fragile]
\frametitle{Simulation 1: Full factorial analysis (vary both)}

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<3way, fig=TRUE, echo=FALSE, height=4, width=8>>=
xyplot(pval_z1 ~ cat | xi, 
       groups = interaction(bin, res_scores), 
       data = d3way, 
       ylim = c(0,0.5),
       type = c("a"), 
       col = c("red", "blue", "red", "blue"),
       lty = c(1,1,2,2),
       xlab = "Categorization",
       ylab = expression(p~"-value of"~Z[1]),
       #index.cond = list(c(1,2,3,4)),
       strip = strip.custom(factor.levels = c(expression(xi==0~"and"~delta==0.3), 
                                              expression(xi==0.8~"and"~delta==1))),
       key = list(text = list(levels(interaction(d3way$bin, d3way$res_scores))),
                  lines = list(lty = c(1,1,2,2), 
                               type = "l",
                               col = c("red", "blue", "red", "blue"))))
@
\end{figure}
\end{frame}

@@

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<tree_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
subdata <- subset(sim_tree, test %in% c("ctree","mfluc",
                                        "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree","MOB",
                                  "GUIDE_scores", "GUIDE"))

@@

xyplot(ari ~ delta | xi, groups = ~ test, 
       data = subdata, 
       type = "l",
       lty = c(1,1,1,1),
       key =  list(text = list(levels(subdata$test)),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal_results, lwd = 1)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8))),
       par.settings = par.settings)

@@

xyplot(ari ~ delta | xi, groups = ~ test, 
       data = prex, 
       type = "l", lty = c(1,1,1,1,2,2,2,2),
       key =  list(text = list(levels(prex$test)[1:4]),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal_results, lwd = 1)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8))),
       par.settings = par.settings
)
@
\end{figure}

@@

Based on the simulation results the following general conclusions can be drawn:
\begin{itemize}
\item Including full model scores is advantageous over using only residuals.
\item Binarizing scores/residuals does not lead to any improvements in performance.
\item Categorizing split variables can simplify calculations, however, important
effects might be missed unless a binning point is close to the true split point.
\item If the applied statistical test performs well as significance test pre-pruning
is preferable. Otherwise post-pruning can compensate this deficiency at least partly.
\end{itemize}
\end{frame}

@@

\footnotesize
Schlosser L, Hothorn T, Zeileis A (2019).
 \dquote{The Power of Unbiased Recursive Partitioning: A Unifying View of CTree, MOB, and GUIDE.}
 \emph{arXiv ...}, arXiv.org E-Print Archive.
 \doi{...}

>>>>

## HCL palette
pal_dgp_light <- qualitative_hcl(3, "Set 2")
pal_dgp <- qualitative_hcl(3, "Dark 3")

@@

{\bf General testing strategy:}
\begin{enumerate}
\item Evaluate a discrepancy measure capturing the observation-wise goodness of fit of $\mathcal{M}(Y,X;\hat{\beta})$.
\item Apply a statistical test assessing dependency of the discrepancy measure to each possible split variable $Z_j$.
\item Select the split variable $Z_j^\ast$ showing the smallest $p$-value.
\end{enumerate}

@@

\begin{frame}
\frametitle{A Unifying View}

The testing strategies of CTree, MOB, and GUIDE can be obtained from a\\
general framework by combining building blocks as follows:

@@

\begin{table}[t]
\begin{tabular}{lllll}
\hline 
      & Scores       & Binarization  & Categorization  & Statistic \\ 
\hline
CTree & Model scores & --	           & -- 	           & Sum of squares\\
MOB   & Model scores & --	           & -- 	           & Maximally selected\\
GUIDE & Residuals    & \checkmark    & \checkmark      & Sum of squares\\
\hline
\end{tabular}
\end{table}

\vspace{0.5cm}
\textit{Note:}\\
\begin{itemize}\item Adaptions of standard versions are provided.\\
\item Further differences (e.g., approximation of null distribution) remain.
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item {\only<4->{\color{lightgray}} Residuals vs.\ full model scores.}
\item {\only<1-3,6->{\color{lightgray}} Binarization of residuals/scores.}
\item {\only<1-5>{\color{lightgray}} Categorization of possible split variables.}
\end{itemize}

@@

\only<2>{
\small
$$
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) \cdot X_1}\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) \cdot X_2}\\
\vdots & {\color{lightgray} \vdots} \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) \cdot X_n}
\end{pmatrix}
%\begin{pmatrix}
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})}\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})}\\
%\vdots & {\color{lightgray} \vdots} \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})}
%\end{pmatrix}
$$
}
\only<3>{
\small
$$
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) \cdot X_1\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) \cdot X_2\\
\vdots & \vdots \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) \cdot X_n
\end{pmatrix}
%\begin{pmatrix} 
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
%\vdots & \vdots \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
%\end{pmatrix}
$$
}
\only<4-5>{
\small
$$
r(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
\begin{pmatrix}
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
\vdots \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
\end{pmatrix}
%
\visible<5>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
> 0\\
\le 0\\
\vdots \\
> 0
\end{pmatrix}
}
$$
}
\only<6-7>{
\small
$$
Z_j = 
\begin{pmatrix}
Z_{j1}\\
Z_{j2}\\
\vdots \\
Z_{jn}\\
\end{pmatrix}
%
\visible<7>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
\text{Q}3\\
\text{Q}1\\
\vdots \\
\text{Q}2
\end{pmatrix}
}
$$
}
\end{frame}

@@

tp <- function (obj, digits = 3, abbreviate = FALSE, 
                fill = c("lightgray", "white"), id = TRUE, 
                just = c("center", "top"), top = 0.85, 
                align = c("center", "left", "right"), gp = NULL, FUN = NULL, 
                height = NULL, width = NULL) 
{
  nam <- names(obj)
  extract_label <- function(node) formatinfo_node(node, FUN = FUN, 
                                                  default = c("terminal", "node"))
  maxstr <- function(node) {
    lab <- extract_label(node)
    klab <- if (is.terminal(node)) 
      ""
    else unlist(lapply(kids_node(node), maxstr))
    lab <- c(lab, klab)
    lab <- try(unlist(lapply(lab, function(x) strsplit(x, 
                                                       "\n"))), silent = TRUE)
    if (inherits(lab, "try-error")) {
      paste(rep("a", 9L), collapse = "")
    }
    else {
      return(lab[which.max(nchar(lab))])
    }
  }
  nstr <- if (is.null(width)) 
    maxstr(node_party(obj))
  else paste(rep("a", width), collapse = "")
  just <- match.arg(just[1L], c("center", "centre", "top"))
  if (just == "centre") 
    just <- "center"
  align <- match.arg(align[1L], c("center", "centre", "left", 
                                  "right"))
  if (align == "centre") 
    align <- "center"
  rval <- function(node) {
    if(node$id == 2) fill <- c("white", pal_dgp_light[1])
    if(node$id == 3 | node$id == 4) fill <- c("white", pal_dgp_light[2])
    if(node$id == 5) fill <- c("white", pal_dgp_light[3])
    fill <- rep(fill, length.out = 2)
    lab <- extract_label(node)
    if (!is.null(gp)) {
      outer_vp <- viewport(gp = gp)
      pushViewport(outer_vp)
    }
    if (is.null(height)) 
      height <- length(lab) + 1L
    node_vp <- viewport(x = unit(0.5, "npc"), y = unit(if (just == 
                                                           "top") 
      top
      else 0.5, "npc"), just = c("center", just), width = unit(1, 
                                                               "strwidth", nstr) * 1.1, height = unit(height, "lines"), 
      name = paste("node_terminal", id_node(node), sep = ""), 
      gp = if (is.null(gp)) 
        gpar()
      else gp)
    pushViewport(node_vp)
    grid.rect(gp = gpar(fill = fill[1]))
    for (i in seq_along(lab)) grid.text(x = switch(align, 
                                                   center = unit(0.5, "npc"), left = unit(1, "strwidth", 
                                                                                          "a"), right = unit(1, "npc") - unit(1, "strwidth", 
                                                                                                                              "a")), y = unit(length(lab) - i + 1, "lines"), 
                                        lab[i], just = align)
    if (id) {
      nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                                                          "npc"), width = max(unit(1, "lines"), unit(1.3, 
                                                                                                     "strwidth", nam[id_node(node)])), height = max(unit(1, 
                                                                                                                                                         "lines"), unit(1.3, "strheight", nam[id_node(node)])))
      pushViewport(nodeIDvp)
      grid.rect(gp = gpar(fill = fill[2], lty = "solid"))
      grid.text(nam[id_node(node)])
      popViewport()
    }
    if (is.null(gp)) 
      upViewport()
    else upViewport(2)
  }
  return(rval)
}

@@

               strip =strip.custom(factor.levels = c(expression("vary"~beta[0]~"and"~beta[1]), 
                                                     expression("vary"~beta[0]), 
                                                     expression("vary"~beta[1])),
                                   bg = "gray90"),
               strip.left=strip.custom(factor.levels = c(expression(xi == 0),                                                              expression(xi == 0.8)),
                                       bg = "gray90"))

@@

%\begin{frame}%[fragile]
%\frametitle{Simulation 1: Full factorial analysis (vary both)}

%\begin{figure}%[t]
%\setkeys{Gin}{width=0.8\linewidth}
%<<3way, fig=TRUE, echo=FALSE, height=4, width=8>>=
%xyplot(pval_z1 ~ cat | xi, 
%       groups = interaction(bin, res_scores), 
%       data = d3way, 
%       ylim = c(0,0.5),
%       type = c("a"), 
%       col = c("red", "blue", "red", "blue"),
%       lty = c(1,1,2,2),
%       xlab = "Categorization",
%       ylab = expression(p~"-value of"~Z[1]),
%       #index.cond = list(c(1,2,3,4)),
%       strip = strip.custom(factor.levels = c(expression(xi==0~"and"~delta==0.3), 
%                                              expression(xi==0.8~"and"~delta==1))),
%       key = list(text = list(levels(interaction(d3way$bin, d3way$res_scores))),
%                  lines = list(lty = c(1,1,2,2), 
%                               type = "l",
%                               col = c("red", "blue", "red", "blue"))))
%@
%\end{figure}
%\end{frame}

@@

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<tree_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
subdata <- subset(sim_tree, test %in% c("ctree","mfluc",
                                        "guide_sum_12", "guide_sum_1_cor"))
#subdata <- subset(subdata, xi %in% c(0,0.5,0.8))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree","MOB",
                                  "GUIDE_scores", "GUIDE"))

@@

xyplot(ari ~ delta | xi, groups = ~ test, 
       data = subdata, 
       type = "l",
       lty = c(1,1,1,1),
       key =  list(text = list(levels(subdata$test)),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal_results, lwd = 1)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8)),
                            bg = "gray90"),
       par.settings = par.settings)

@@

#prex <- subset(prex, xi %in% c(0,0.5,0.8))

xyplot(ari ~ delta | xi, groups = ~ test, 
       data = prex, 
       type = "l", lty = c(1,1,1,1,2,2,2,2),
       key =  list(text = list(levels(prex$test)[1:4]),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal_results, lwd = 1)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8)),
                            bg = "gray90"),
       par.settings = par.settings
)
@
\end{figure}

@@

\vspace{0.4cm}

In this setting:

\vspace{0.2cm}

\begin{itemize}
\setlength\itemsep{0.2cm}
\item Include full model scores better than residuals only.
\item Use original values of scores/residuals better than binarized values.
\item Categorization is simpler, but less powerful in margins.
\item If the significance test performs well pre-pruning works well,\\
otherwise post-pruning might be needed.
%\item Including full model scores is advantageous over using only residuals.
%\item Binarizing scores/residuals does not improve the performance.
%\item Categorizing split variables can simplify calculations, however,
%effects might be missed unless a binning point is close to the true split point.
%\item If the applied test performs well as significance test pre-pruning
%is preferable, otherwise post-pruning.
\end{itemize}
\end{frame}

@@

\footnotesize
Schlosser L, Hothorn T, Zeileis A (2019).
 \dquote{The Power of Unbiased Recursive Partitioning: A Unifying View of CTree, MOB, and GUIDE.}
 \emph{paper in preparation}.
